Carbon cathodes for rechargeable aluminium-ion batteries

### 2020-03-17 manuscript.tex

\begin{document}
\maketitle
\begin{abstract}
 In this work, we compare four different forms of carbon: activated carbon from human hair, activated carbon from hemp fibers, a fullerene mixture consisting of \ce{C60} and \ce{C70} fullerenes (CFEx) and Super-P carbon black (SPCB) as cathodes for non-aqueous aluminium-ion batteries. These materials differ in their general structure, porosity and morphology. The fullerenes display a crystalline structure, whereas hemp fibers, SPCB and hair are amorphous in nature. Of all materials, activated carbon from human hair recorded the highest specific capacity after 50 cycles at 103 mAh g$^{-1}$ with a coulombic efficiency of $\sim$90\% at a current rate of 50 mV s$^{-1}$. Both hemp fibers and SPCB achieved their highest specific capacities at 56 mAh g-1 and 84 mAh g$^{-1}$ respectively. CFEx recorded its highest capacity at 78 mAh g$^{-1}$ and maintained it for 50 cycles. The cells were charged and discharged to 2.45 V and 0.2 V respectively. 
 \end{abstract}

@@

\section{Introduction}
Different varieties of carbon-based materials have been widely used in energy storage applications. Graphite, with its layered structure turned out to be an ideal intercalation cathode material \cite{ji_recent_2011, yoo_large_2008, lian_large_2010}. Activated carbon, owing to its porous structure, provides a high surface area for absorption of electrolyte ions in super-capacitors \cite{eliad_ion_2001, zhu_carbon-based_2011-2}. Non-aqueous aluminium-ion batteries (AIBs) use low-cost, abundant materials, a non-flammable electrolyte and provide a higher theoretical energy density than lithium-ion batteries (LIBs) due to the multivalent nature of aluminium. They offer an interesting potential long-term alternative to the problematic LIB technology \cite{ambroz_trends_2017-1}. In this work, we tested a number of rechargeable AIBs using an ionic liquid electrolyte with different carbon-based cathodes. The aim of this study was to systematically explore the properties of different carbon-materials when used as AIB cathodes.

Graphite has been repeatedly used as a cathode in AIBs. It has (i) a layered structure that enhances the intercalation process, (ii) good conductivity, and (iii) a high electrical potential {\it vs.} \ce{Al}/\ce{Al^3+} of 2.1 V. Various forms of graphite such as fluorinated graphite \cite{rani_fluorinated_2013}, kish graphite flakes \cite{wang_kish_2017}, three-dimensional (3D) graphitic-foam\cite{wu_3d_2016}, graphene aerogels\cite{huang_graphene_2019} and several other forms have been tested as cathodes for AIBs, showing typical specific capacities ranging from 60-250 mAh g$^{-1}$. The \ce{AlCl4-}-anions intercalate into the graphitic stacks when the cell is being charged and deintercalate during discharge. X-ray diffraction (XRD) and Raman spectroscopy studies helped in establishing this mechanism \cite{rani_fluorinated_2013, wang_advanced_2017,lin_ultrafast_2015-3} as shown in Figure \ref{fig:graphitemech}. For example: a Raman spectra showed conversion of a doublet peak during charge to one single peak after the cell was fully charged, indicating two stages of intercalation. XPS studies confirmed reversible oxidation/reduction of carbon when \ce{AlCl4-} anions intercalate/deintercalate respectively \cite{stadie_zeolite-templated_2017, liu_binder-free_2019,wei_amorphous_2017}.

\begin{figure}[h!]
  \centering
  \includegraphics[width=\textwidth]{fig/graphitemech}
    \caption{Intercalation of \ce{AlCl4-} ions during charging and deintercalation during discharging in an Al/graphite cell. The interlayer distance between two graphite sheets is 3.3 \AA.}
  \label{fig:graphitemech}
\end{figure}

In this study, we compared four carbon-based materials with carbonised natural products. Super-P has been the premium choice for a carbon additive, while preparing a slurry, as it improves the conductivity of the active material. Using it as an active material, we tried to find if Super-P adds any capacity of its own to the battery's specific capacity. Activated carbon from natural products like rice or wood have been previously used in super-capacitors \cite{hussain_development_2019, frackowiak_carbon_2001}. Also known as 'hierarchical porous carbons', these structures contain pores of various sizes (mesopores and micropores), which enhance the surface area of the electrodes. The pores absorb the charge carriers reversibly on their surface by faradaic/ non-faradaic processes. Based on this hypothesis, we tested activated carbon from human hair and hemp fibers as cathodes for non-aqueous AIBs.

\section{Results and discussion}

\begin{table}[h!]
\caption{Comparing battery metrics of all carbon-based cathodes tested in this work} \label{table1}
\begin{center}
\begin{tabular}{|lcccc|}
\hline
Active material & {\textbf{Size}} & {\textbf{Specific capacity}} & {\textbf{Cell efficiency}} & {\textbf{Cell voltage}}\\
 & {\textbf{(pore size)}} & {\textbf{(mAh g$^{-1}$)}} & {\textbf{(\%)}} & {\textbf{(V)}}\\
\hline
Human hair & 5${\mu}$ m & 102 & 97 & 1.9 \\
Fullerene mix & 8.8 \AA & 78 & 85 & 1.7 \\
Hemp fibers & 2.3 $\mu$ m & 49 & 75 & 1.8 \\
SPCB & 300 \AA & 46 & 40 & 1.5 \\
\hline  % Please only put a hline at the end of the table
\end{tabular}
\end{center}
\end{table}

\begin{figure}[h!]
  \centering
  \includegraphics[width=\textwidth]{fig/CDCall}
    \caption{Specific capacities of AC (human hair, hemp fibers), CFEX and SPCB in their a) first and b) 50$^{th}$ cycle at a current rate of 50 mA g$^{-1}$. c) Coulombic efficiencies of cells at a current rate of 50 mAg$^{-1}$. d) Galvanostatic charge/discharge profile of all cells at various current rates ranging from 25 mAg$^{-1}$ to 100 mAg$^{-1}$ in a two-electrode setup against Al$^{3+}$/Al.}
  \label{fig:CDCall}
\end{figure}

The battery system comprised a cathode, 99\% pure aluminium foil as the anode, and a room temperature ionic liquid (RTIL) as the electrolyte. Activated carbon derived from human hair and hemp fibers, and SPCB have a non-crystalline structure, however their Raman spectra revealed the presence of a few graphitic planes. Graphitic layers would allow \ce{AlCl4-} ions to intercalate during charging. CFEx (a mixture of \ce{C60} and \ce{C70} fullerenes) does not have a layered structure. Both fullerenes have a cage-like morphology and the chloroaluminates are not small enough to move in and out of them during charge/discharge. We suggest that \ce{AlCl4-} anions migrated through the gaps present in between the fullerenes without altering its molecular structure. Using a \textit{Neware BTS 3000} battery analyser, specific capacities of the cathodes were recorded at a constant current density of 50 mA g$^{-1}$ (cf.\ Figure \ref{fig:CDCall}a and b). Morphology of the cathodes before and after the cycles were studied using Raman spectroscopy, X-ray diffraction (XRD) patterns and scanning electron microscopy (SEM).

Figure \ref{fig:CDCall}a and b compares the specific capacities of all cells for their first and 50$^{th}$ cycles. Human hair cathodes exhibited a high capacity of $\sim$100 mAh g$^{-1}$ with coulombic efficiency of $\sim$95$\%$. High surface area due to its porous structure might have led to an additional 'capacitor-like' charge storage \cite{frackowiak_carbon_2001}. Hemp batteries displayed a capacity of 56 mAh g$^{-1}$ in their first cycle, which decreased to 45 mAh g$^{-1}$ after 50 cycles. CFEx displayed a capacity of around 80 mAh g$^{-1}$ with coulombic efficiency of $\sim$90\%. With an initial value of 84 mAh g$^{-1}$, specific capacity of SPCB decreased to 47 mAh g$^{-1}$ with a low coulombic efficiency of $\sim$40\%. Capacity for SPCB and hemp cells decreased considerably after repeated charge/discharge cycles. A low coulombic efficiency (observed in both hemp and SPCB) can be attributed to side reactions in a battery. These may include electrode or electrolyte interactions with impurities, or degradation of the cathode structure (pulverisation) \cite{gyenes_understanding_2015-1}. However, capacity fade for CFEx and human hair cells was minimal. This suggests that CFEx and human hair have a more stable structure and have the potential to store charge reversibly \cite{pramanick_human_2016}.

% In the above paragraph, you discuss 3 specific cells. This is suicidal (metaphorically speaking). In science you make a point by showing that something is reproducible. Having done just 3 expriments is asking for the reviewer to knock it down. You need to re-write this in a more general manner.
\fbox{\begin{minipage}{46.5em}
\section*{Activation of carbon}
The production of activated carbon consists of carbonisation of a precursor at a temperature below 900$^{\circ}$ C in an inert atmosphere and chemical or physical activation of the carbonised precursor. Activating agents play an important role in determining the porosity of an activated carbon \cite{arenas_effect_2004}. Using alkali hydroxides at high temperature creates micropores which increases the surface area of the material \cite{dong_commercial_2019, liu_hair-based_2017}. In this work, sodium hydroxide (NaOH) was used as the activating agent. The reaction that takes place inside the carbon matrix after adding NaOH is as follows:

@@

In this reaction, NaOH was reduced to free metal, Na. These atoms in turn expanded the carbon matrix after intercalating into the carbon structure. Increased temperature (750$^{\circ}$ C) forced the  atoms out of the carbon matrix, thus creating micropores. Oxidation of carbon from oxygen atoms of the hydroxide group formed carbon dioxide (\ce{CO2}), providing routes for channeling the sodium atoms into the internal structure, resulting in a well-connected porous structure \cite{satish_macroporous_2015}. The calcinating temperature used here was 750$^{\circ}$ C. Figure \ref{fig:ACHsyn} displays a flowchart describing activated carbon synthesis. 
\end{minipage}}

\begin{figure}[ht!]
\centering
\includegraphics[width=\textwidth]{fig/ACHsyn}
\caption{Synthesis of activated carbon (AC) from human hair using NaOH as the activating agent.}
\label{fig:ACHsyn}
\end{figure}

% Paragraph above: you need to prove/reason every statement that you make. You just claim a lot of stuff without providing any evidence. Also: ARTICLES!!!!!

%Unfortunately the structural domains appeared to have been damaged after this process. This phenomena negatively impacted its charge-storing capacity.

% Above paragraph: Lots of unnecerssary information (see comments). Focus on what is relevant. Also sentences like the Carbon Valley one are suicidal. When you write, always have your worst scientific enemy in mind and write it so that you win.
\begin{figure}[h!]
\centering
\includegraphics[width=0.75\textwidth]{fig/CFExsol}
\caption{Comparison of solubility of a) pure \ce{C60} fullerene b) fullerene extract (CFEx) and AC derived from hemp fibers in \ce{AlCl3}/EMImCl ionic liquid electrolyte.}
\label{fig:CFExsol}
\end{figure}

Fullerenes have a fused-ring structure with a nucleus-to-nucleus diameter of 7.1\AA\ and a van der Waals (vdW) diameter of 11\AA\ in a single crystal. However, they are zero-dimensional materials, which means they cannot provide an efficient path for electron transport or a long-range conductivity \cite{loutfy_fullerene_2002, winkler_two-component_2007}. They are known to be weak battery materials owing to their solubility in electrolytes, especially in LIBs \cite{seger_prospects_1991}. To test their solubility in \ce{AlCl3}-EMImCl, 100 mg of CFEx was mixed in the electrolyte and stirred for 24 hours. The solution was left to stand for another 24 hours inside a \ce{N2}-filled glove box. CFEx seemed to dissolve in the RTIL (Figure \ref{fig:CFExsol}) since no phase separation was observed. It has been reported that poly-sulphides (formed during charge/discharge cycles) are soluble in the electrolyte of a Li-S battery. They form an insulating layer of \ce{Li2S} on the anode, which results in capacity fading \cite{sun_effect_2017}. Since no such effect was observed in aluminium-fullerene cells, the solubility of fullerenes does not seem to impact the charge-storing capacity of CFEx. On the contrary, aluminium-fullerene batteries demonstrated excellent capacity retention at various current rates(Figure \ref{fig:CDCall}d).

 \begin{figure}[h!]
  \centering
  \includegraphics[width=\textwidth]{fig/superPmech}
    \caption{Suggested mechanism for an aluminium-SPCB cell. \ce{AlCl4-} anions intercalate into the very few graphitic planes in Super-P, while few anions adsorb onto its surface. However, further cycling leads to cathode pulverisation, which results in capacity fading.}
  \label{fig:superPmech}
\end{figure}

Super-P is an amorphous form of carbon with a high surface area of 62 m$^2$ g$^{-1}$ and a highly disordered structure. The pore sizes range from $\sim$30-50 nm \cite{younesi_analysis_2015}. As a cathode material, SPCB underwent the highest capacity loss within 30 cycles (45\%). It seems continuous cycling destroyed the structural arrangement of the carbon atoms, which resulted in further alleviated capacity and low coulombic efficiencies.

 \begin{figure}[h!]
  \centering
  \includegraphics[width=\textwidth]{fig/hempmech}
    \caption{Suggested mechanism for an aluminum-hemp cell. Hemp fibers have pore sizes which might be as large as 2.0-2.5 $\mu$m allowing the \ce{AlCl4-} to get absorbed on their surface. However, agglomeration of these fibers after a few cycles reduces the number of active sites available for effective charge storage.}
  \label{fig:hempmech}
\end{figure}

 \begin{figure}[h!]
  \centering
  \includegraphics[width=\textwidth]{fig/raman}
    \caption{Raman spectra of pristine (in black) and charged (in red) a) CFEx, b) hemp, c) human hair and d) SPCB cathodes displaying presence of both D and G bands.}
  \label{fig:raman}
\end{figure}

It has been previously established that chloroaluminates intercalate into the graphitic planes during charging \cite{lin_ultrafast_2015-3}. Since activated carbon and SPCB cathodes tested in this work have graphitic planes present in them, we compared the pristine and charged electrodes only. Figure \ref{fig:raman} displays the Raman spectra of the pristine and charged (intercalated) cathodes. Pristine hair, hemp and SPCB cathodes had a significant D-band present in them, indicating absence of an ordered structure (at $\sim$1300 cm$^{-1}$ for human hair, 1329.7 cm$^{-1}$ for hemp fibers, and 1352.0 cm$^{-1}$ for SPCB). Broad peaks at about 1600 cm$^{-1}$ and about 1350 cm$^{-1}$ arise from \ce{sp2} clusters like $\alpha$-carbons with bond angle disorder \cite{shimo}. Raman spectra of CFEx in Figure \ref{fig:raman}c displayed the characteristic features of both \ce{C60} and \ce{C70} molecules. A 'pentagonal pinch mode', usually observed in a \ce{C60} Raman spectrum, was present at 1460 cm$^{-1}$. \ce{C70} has multiple bands because it has a reduced molecular symmetry, which increases the number of vibrational modes, consequently increasing the number of active Raman bands \cite{kimbrell_analysis_2014}. It was interesting to note that the spectra of charged CFEx looked strikingly similar to the pristine ones, Figure \ref{fig:raman}c. Since Raman spectroscopy is sensitive to minute differences in the molecular morphology, results suggested that the fullerenes did not undergo any significant structural change during cycles. Charged AC (from hair) cathodes showed an increased full-width at half maxima (FWHM), which suggested an increase in the lattice defects and structure deformities. Obviously, these defects arise after intercalation of chloroaluminates into the few graphitic planes present in the carbon matrix. Furthermore, a highly porous structure allows surface adsorption of charge carrying species (much like in super-capacitors) \cite{beguin_carbons_2014}. Intercalation into graphitic planes/micropores, along with surface adsorption of chloroaluminates, result in aluminium-hair batteries performing better than other materials by storing more charge \cite{brezesinski_ordered_2010}. Schematic of an aluminium-hair battery is illustrated in Figure \ref{fig:ACHmech}.

 \begin{figure}[h!]
  \centering
  \includegraphics[width=\textwidth]{fig/ACHmech}
    \caption{Suggested mechanism for an aluminium-human hair cell. Chloroaluminates (\ce{AlCl4-}) intercalate into the few graphitic planes and micro/mesopores present in them, in addition to surface adsorption of ions displaying both Faradaic and non-Faradaic processes for charge storage.}
  \label{fig:ACHmech}
\end{figure}

%contains about 51\% carbon, 17\% nitrogen, 21\% oxygen, 6\% hydrogen, 5\% sulfur, and trace amounts of iron, magnesium, arsenic, chromium and various minerals \cite{lee_chemistry}.  
Hemp fibers and SPCB had a highly disordered structure to begin with. Repeated intercalation or absorption of ions on their surface in the first few cycles would have further increased the lattice defects, which is one of the reasons why hemp and SPCB batteries failed to retain their capacity. However, CFEx does not have graphitic planes to intercalate chloroaluminate ions. Since the fullerenes have a very high surface area, surface adsorption of ions is highly likely \cite{adams_van_1994}. Furthermore, it might be possible for the anions to seep through the gaps present in between two fullerenes (as shown in Figure \ref{fig:CFExmech}) and for the surface-based redox processes to take place in a systematic way.

 \begin{figure}[tbh!]
  \centering
  \includegraphics[width=\textwidth]{fig/CFExmech}
    \caption{Suggested mechanism for a fullerene cell. \ce{AlCl4-} ions seeping in through the gaps present between fullerene molecules.}
  \label{fig:CFExmech}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.8\textwidth]{fig/SEM}
    \caption{Scanning electron microscopy (SEM) images comparing pristine a) human hair, b) hemp, c) CFEx and d) SPCB and charged e)human hair, f) hemp, g) CFEx and h) SPCB cathodes. Hemp fibers and SPCB undergo permanent changes after charge/discharge cycles and fail to retain capacity.}
  \label{fig:SEM}
\end{figure}

Figure \ref{fig:SEM} shows the SEM images of pristine (Figure \ref{fig:SEM}a, b, c and d) and charged (\ref{fig:SEM}e,f,g and h) cathodes of all carbon materials. Activated carbon has a highly porous structure as seen in Figure \ref{fig:SEM}b and d). However, hemp cathodes lost their porosity and degraded after 30 cycles, Figure \ref{fig:SEM}f). Significant agglomeration was observed in SPCB as well (Figure \ref{fig:SEM}d) and h). It was interesting to note that the cycled AC (from hair) and fullerene cathodes did not undergo any significant changes after cycles.

>>>>

\begin{document}
\maketitle
\begin{abstract}
 In this work, four different forms of carbon: activated carbon (AC) from human hair, AC from hemp fibers, a carbon fullerene extract consisting of \ce{C60} and \ce{C70} fullerene (CFEx) and Super-P carbon black (SPCB) were tested and compared as cathodes for non-aqueous aluminium-ion batteries (AIBs). These materials differ in their general structure, porosity and morphology. The fullerenes display a crystalline structure, whereas hemp fibers, SPCB and hair are amorphous in nature. Of all materials, AC obtained from human hair recorded the highest specific capacity after 50 cycles at 103 mAh g$^{-1}$ with a Coulombic efficiency (CE) of $\sim$90\% at a current rate of 50 mV s$^{-1}$. Both hemp fibers and SPCB achieved their highest specific capacities at 56 mAh g-1 and 84 mAh g$^{-1}$ respectively. CFEx recorded its highest capacity at 78 mAh g$^{-1}$ and maintained it for 50 cycles. The cells were charged and discharged to 2.45 V and 0.2 V respectively. 
 \end{abstract}

@@

\section{Introduction}
Non-aqueous aluminium-ion batteries (AIBs) use low-cost, abundant materials, a non-flammable electrolyte and provide a higher theoretical energy density than lithium-ion batteries (LIBs) due to the multivalent nature of aluminium (Al) \cite{lin_ultrafast_2015-2,paranthaman_transformational_2010,wang_advanced_2017}. They offer an interesting potential long-term alternative to the LIB technology \cite{ambroz_trends_2017-1}. Different varieties of carbon-based materials have been widely used in various energy storage applications. Graphite, with its layered structure turned out to be an ideal intercalation cathode material for many types of ions \cite{ji_recent_2011, yoo_large_2008, lian_large_2010}. It allows reversible intercalation of ions in its layers during charging and discharging processes, which results in high specific capacities. Owing to its porous structure, activated carbon (AC) provides a high surface area, and is a typical electrode material used in super-capacitors \cite{eliad_ion_2001, zhu_carbon-based_2011-2}. In this work, a number of rechargeable AIBs were tested using an ionic liquid electrolyte with different carbon-based cathodes. The aim of this study was to systematically explore the properties of different carbon materials when used as AIB cathodes.

Graphite has been repeatedly used as a cathode in AIBs \cite{lin_ultrafast_2015-2, wang_advanced_2017,zhang_novel_2016-1, rani_fluorinated_2013} due to its layered structure that enhances the intercalation process, good conductivity, and high electrical potential {\it vs.} \ce{Al}/\ce{Al^3+} of 2.1 V. Various forms of graphite such as fluorinated graphite \cite{rani_fluorinated_2013}, kish graphite flakes \cite{wang_kish_2017}, three-dimensional (3D) graphitic-foam \cite{wu_3d_2016}, graphene aerogels \cite{huang_graphene_2019} and several other forms have been tested as cathodes for AIBs, showing specific capacities ranging from 60--250 mAh g$^{-1}$. As per the established mechanism, \ce{AlCl4-}-anions intercalate into graphite layers when the cell is being charged and deintercalate during discharge. X-ray diffraction (XRD) and Raman spectroscopy studies helped in establishing this mechanism \cite{rani_fluorinated_2013, wang_advanced_2017,lin_ultrafast_2015-2} as shown in Figure \ref{fig:graphitemech}. For example, Raman spectra showed conversion of a doublet peak during charge to one single peak after the cell was fully charged, indicating two stages of intercalation \cite{wang_advanced_2017}. X-ray photoelectron spectroscopy (XPS) studies confirmed reversible oxidation/reduction of carbon when \ce{AlCl4-} anions intercalate/deintercalate respectively \cite{stadie_zeolite-templated_2017, liu_binder-free_2019,wei_amorphous_2017}.

\begin{figure}[] %[h!]
  \centering
  \includegraphics[width=\textwidth]{fig/graphitemech}
    \caption{Intercalation of \ce{AlCl4-} ions during charging and deintercalation during discharging in an Al/graphite cell. The interlayer distance between two graphite sheets is 3.3 \AA.}
  \label{fig:graphitemech}
\end{figure}

In this work, four different carbon-based materials --- Ac from human hair, AC from hemp fibers, fullerene extract (CFEx) and Super-P carbon black (Super-P) were investigated as cathodes for AIBs. Super-P is an amorphous form of carbon. It is highly conductive and is added in electrode slurries to enhance the conductivity of a cathode material. Activated carbon derived from natural products such as rice husk, coconut shells or wood, have been previously used in batteries and super-capacitors \cite{hussain_development_2019, frackowiak_carbon_2001}. Most of the carbon-based materials have a high surface area. AC derived from both hair and hemp fibers contain pores of various sizes (mesopores and micropores). Fullerenes on the other hand, has a tightly packed structure with Brunauer-Emmet and Teller (BET) surface area of 19 m$^2$ g$^{-1}$ \cite{}.

\section{Results and discussion}
The battery system comprised a cathode, 99\% pure Al foil as the anode, and a room temperature ionic liquid (RTIL) as the electrolyte. Super-P is an amorphous form of carbon with a high surface area of 62 m$^2$ g$^{-1}$ and a highly disordered structure \cite{see_reversible_2017}. The pore sizes range from $\sim$30-50 nm \cite{younesi_analysis_2015}. AC derived from human hair, hemp fibers, and SPCB all have a non-crystalline structure. However their Raman spectra revealed the presence of a few graphitic planes. These planes would allow \ce{AlCl4-} ions to intercalate during charging. CFEx (a mixture of \ce{C60} and \ce{C70} fullerenes) does not have a layered structure. Fullerenes have a cage-like morphology and the chloroaluminates are not small enough to move in and out of them during charge/discharge. We assumed that \ce{AlCl4-} anions migrated through the gaps present in between the fullerenes and interacted with their surface. Specific capacities and Coulombic efficiencies (CEs) of the cathodes were recorded at various current rates (cf.\ Figure \ref{fig:CDCall}a and b). Morphology of the cathodes before and after the cycles were studied using Raman spectroscopy, XRD patterns and scanning electron microscopy (SEM).

\begin{figure}[]
  \centering
  \includegraphics[width=\textwidth]{fig/CDCall}
    \caption{Specific capacities of AC (human hair, hemp fibers), CFEx and SPCB in their a) first and b) 50$^{th}$ cycle at a current rate of 50 mA g$^{-1}$. c) Coulombic efficiencies (CEs) of cells at a current rate of 50 mAg$^{-1}$. d) Galvanostatic charge/discharge profile of all cells at various current rates ranging from 25 mAg$^{-1}$ to 100 mAg$^{-1}$ in a two-electrode setup against Al$^{3+}$/Al.}
  \label{fig:CDCall}
\end{figure}

Figures \ref{fig:CDCall}a) and b) show the specific capacities of all cells for their first and 50$^{th}$ cycles. Human hair cathodes exhibited a high capacity of $\sim$100 mAh g$^{-1}$ with CE of $\sim$95\% shown in Figure \ref{fig:CDCall}c). Hemp batteries displayed a capacity of 56 mAh g$^{-1}$ in their first cycle, which decreased to 45 mAh g$^{-1}$ after 50 cycles. CFEx displayed a capacity of around 80 mAh g$^{-1}$ with CE of $\sim$90\% and maintained that for 50 cycles. With an initial value of 84 mAh g$^{-1}$, specific capacity of Super-P decreased to 47 mAh g$^{-1}$ and a low CE of $\sim$40\% was observed. The discharging capacity of SPCB and hemp fibers decreased considerably after repeated charge/discharge cycles. A low CE that was observed in both hemp and Super-P, can be attributed to side reactions in the batteries. These may include electrode or electrolyte interactions with impurities or degradation of the cathode structure (pulverisation) \cite{gyenes_understanding_2015-1}. The capacity loss for CFEx and human hair cells was minimal. This suggests that CFEx and human hair have a more stable structure and have the potential to store charge reversibly \cite{pramanick_human_2016}.

\section*{Activation of carbon}
The production of AC consists of carbonisation of a precursor at a temperature below 900$^{\circ}$ C in an inert atmosphere and chemical or physical activation of the carbonised precursor. Activating agents play an important role in determining the porosity of AC \cite{arenas_effect_2004}. Using alkali hydroxides at high temperature creates micropores, which increases the surface area of the material \cite{dong_commercial_2019, liu_hair-based_2017}. In this work, sodium hydroxide (NaOH) was used as the activating agent. The following reaction is taking place within the carbon matrix after addition of NaOH:

@@

\begin{figure}%[ht!]
\centering
\includegraphics[width=\textwidth]{fig/ACHsyn}
\caption{Synthesis of activated carbon (AC) from human hair using NaOH as the activating agent.}
\label{fig:ACHsyn}
\end{figure}

% Paragraph above: you need to prove/reason every statement that you make. You just claim a lot of stuff without providing any evidence. Also: ARTICLES!!!!!
Fullerenes have a fused-ring structure with a nucleus-to-nucleus diameter of 7.1\AA\ and a van der Waals (vdW) diameter of 11\AA\ in a single crystal. However, they are zero-dimensional materials, which means they cannot provide an efficient path for electron transport or a long-range conductivity \cite{loutfy_fullerene_2002, winkler_two-component_2007}. They are known to be weak battery materials owing to their solubility in electrolytes, especially in LIBs \cite{seger_prospects_1991}. To test their solubility in \ce{AlCl3}-EMImCl, 100 mg of CFEx were mixed in the electrolyte and stirred for 24 hours. The solution was left to stand for another 24 hours inside a \ce{N2}-filled glove box. CFEx seemed to dissolve in the RTIL (Figure \ref{fig:CFExsol}) since no phase separation was observed. It has been reported that poly-sulphides (formed during charge/discharge cycles) are soluble in the electrolyte of a Li-S battery. They form an insulating layer of \ce{Li2S} on the anode, which results in capacity fading \cite{sun_effect_2017}. Since no such effect was observed in the fullerene cells, this phenomena does not seem to impact its charge-storing capacity. On the contrary, Al/CFEx cells demonstrated excellent capacity retention at various current rates (Figure \ref{fig:CDCall}d)).

\begin{figure}
\centering
\includegraphics[width=0.75\textwidth]{fig/CFExsol}
\caption{Comparison of solubility of a) pure \ce{C60} fullerene b) fullerene extract (CFEx) and AC derived from hemp fibers in \ce{AlCl3}/EMImCl ionic liquid electrolyte.}
\label{fig:CFExsol}
\end{figure}

It has been previously established that chloroaluminates intercalate into the graphitic planes during charging \cite{lin_ultrafast_2015-2}. Since both AC and SPCB cathodes contain graphitic planes, we expected intercalation during cell charging. To study these changes, charged cathodes of all materials were analysed by Raman spectroscopy and compared with the pristine ones. The data obtained is displayed in Figure \ref{fig:raman}. Graphite shows a characteristic D-band present at 1300 cm$^{-1}$, which originates from the hybridized vibrational modes associated with graphene edges and indicates the absence of long-range order. Pristine hair, hemp and SPCB cathodes showed a significant D-band present at $\sim$1300 cm$^{-1}$ for human hair, 1329.7 cm$^{-1}$ for hemp fibers, and 1352.0 cm$^{-1}$ for SPCB, indicating the absence of a defined lattice structure. Both the peaks at 1600 cm$^{-1}$ and 1350 cm$^{-1}$ are broad due to the presence of \ce{sp2} clusters like $\alpha$-carbons, which have a bond angle disorder \cite{shimodaira_raman_2002}.

 \begin{figure}
  \centering
  \includegraphics[width=\textwidth]{fig/raman}
    \caption{Raman spectra of pristine (in black) and charged (in red) a) CFEx, b) AC from hemp fibers, c) from human hair (ACH) and d) Super-P cathodes displaying presence of both D and G bands.}
  \label{fig:raman}
\end{figure}

Raman spectra of CFEx in Figure \ref{fig:raman}c) displayed the characteristic bands of both \ce{C60} and \ce{C70} molecules. A 'pentagonal pinch mode', usually observed in a \ce{C60} Raman spectrum, was present at 1460 cm$^{-1}$. It was observed that \ce{C70} had multiple bands. This was due to its reduced molecular symmetry, which increased the number of vibrational modes, consequently increasing the number of active Raman bands \cite{kimbrell_analysis_2014}. It was interesting to note that the spectra of charged CFEx looked almost identical to the pristine ones. The results suggested that the fullerenes did not undergo any significant structural change during the cycles. As a result, the cells displayed a very stable CE. Furthermore, fullerenes stored the same amount of charge after every cycle (similar discharge capacity after 50 cycles) without undergoing any significant structural changes.

Charged AC (from hair) cathodes showed an increased full-width at half maxima (FWHM), which suggested an increase in the lattice defects and structure deformities. These defects would arise if the chloroaluminate ions intercalated into the few graphitic planes present in the carbon matrix. However, a highly porous structure would also allow surface adsorption of charge carrying species much like in super-capacitors \cite{beguin_carbons_2014}. Since activated carbon lacks the long-range ordered structure present in natural graphite, intercalation into short-range graphitic planes, as well as surface adsorption of chloroaluminates, can contribute to its high capacity \cite{brezesinski_ordered_2010}. This might have resulted in aluminium-hair batteries performing better than other materials by storing more charge as a result of the dual mechanism. A schematic of the proposed mechanism of an aluminium-hair battery is displayed in Figure \ref{fig:ACHmech}.

 \begin{figure}%[h!]
  \centering
  \includegraphics[width=\textwidth]{fig/ACHmech}
    \caption{Suggested mechanism for an Al/hair cell. Chloroaluminate ions (\ce{AlCl4-}) intercalate into the few graphitic planes and micro/mesopores present in them, in addition to surface adsorption of ions displaying both Faradaic and non-Faradaic processes for charge storage.}
  \label{fig:ACHmech}
\end{figure}

%contains about 51\% carbon, 17\% nitrogen, 21\% oxygen, 6\% hydrogen, 5\% sulfur, and trace amounts of iron, magnesium, arsenic, chromium and various minerals \cite{lee_chemistry}.

Figure \ref{fig:allmech}b) and c) show the proposed mechanism of charge storage in an Al/hemp and Al/SPCB cell. Both hemp fibers and SPCB had a highly disordered structure to begin with (cf.\ Raman spectra). Repeated intercalation or adsorption of ions onto their surface during the first few cycles further damaged their structure, which is one of the reasons why hemp and SPCB batteries failed to retain their capacity. However, CFEx does not have graphitic planes to intercalate chloroaluminate ions. Since the fullerenes have a very high surface area, surface adsorption of ions is highly likely \cite{adams_van_1994}. Furthermore, it might be possible for the anions to seep through the gaps present in between two fullerenes (\ce{C60} and \ce{C70}) as shown in Figure \ref{fig:allmech}a) and for the surface-based redox processes to take place in a systematic way throughout the electrode.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{fig/allmech}
    \caption{Suggested mechanism for an a) Al-CFEx cell, b) Al-hemp cell, hemp fibers have pore sizes as large as 2.0-2.5 $\mu$m allowing the \ce{AlCl4-} to get absorbed on their surface. However, agglomeration of these fibers after a few cycles reduces the number of active sites available for effective charge storage, and c) Al-Super-P cell, chloroaluminates intercalate into the very few graphitic planes in Super-P, while few anions adsorb onto its surface. However, further cycling leads to cathode pulverisation, which results in capacity fading.}
  \label{fig:allmech}
\end{figure}

Figure \ref{fig:SEM} shows the SEM images of pristine (Figure \ref{fig:SEM}a), b), c) and d)) and charged (\ref{fig:SEM}e) ,f) ,g) and h)) cathodes of all carbon materials. Human hair and hemp fibers have a highly porous structure as seen in Figure \ref{fig:SEM}b) and d). However, hemp lost its surface porosity after 30 cycles, Figure \ref{fig:SEM}f). In addition, Figure \ref{fig:SEM}d) and h) implies that SPCB underwent significant agglomeration during cycles. It was interesting to note that the fullerene cathode retained its morphology after 50 cycles.

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{fig/SEM}
    \caption{Scanning electron microscopy (SEM) images comparing pristine a) human hair, b) hemp, c) CFEx and d) SPCB and charged e) human hair, f) hemp, g) CFEx and h) SPCB cathodes. Hemp fibers and SPCB undergo permanent changes after charge/discharge cycles and fail to retain capacity.}
  \label{fig:SEM}
\end{figure}

Pristine (in black), charged (in green) and discharged (in red) cathodes of CFEx and human hair are showed in Figure \ref{fig:XRD}a) and b) respectively. Figure \ref{fig:XRD}a) displays the characteristic peaks of both \ce{C60} and \ce{C70} at 2$\theta$ values of 10.9$^{\circ}$, 17.8$^{\circ}$, 20.9$^{\circ}$ and 28.2$^{\circ}$ for \ce{C60} and 18.9$^{\circ}$, 19.3$^{\circ}$ and 21.8$^{\circ}$ for \ce{C70} molecules. New diffraction peaks at lower 2$\theta$ values appeared for charged electrodes. However, after discharge, the XRD patterns looked similar to the ones of the pristine cathodes. This data strongly suggests a reversible process taking place during cycling. To confirm this, the unit cell lattice parameters for both pristine and charged cathodes using \ce{C60} were calculated. The unit cell had a tetragonal crystal symmetry with a space group of \textit{P42/ mmc} and a space group number 131 (ICDD: 04-013-1339). Lattice parameters 'a' and 'b' for the charged cathode increased from 9.06 \AA\ to 9.57 \AA\ and 'c' increased from 15.03 \AA\ to 15.65 \AA, as shown in Figure \ref{fig:cfexcrys}a) and b). Lattice parameters of a discharged fullerene were closer to pristine values. These changes suggested a reversible insertion of chloroaluminates into the free spaces between fullerene molecules taking place. A possible site for \ce{AlCl4-} intercalation is depicted in Figure \ref{fig:cfexcrys}c).
